A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-20 docs/icccn2009/paper.tex
Minor edits
git-svn-id: svn://localhost/netbeams/trunk@420 559edde2-a4d4-44e0-96fc-4035e40d7396

Each probe may have multiple sensors.  A computer can interface with
the Sonde via RS-232C or SDI-12. The Sonde can be configured to
collect data samples in discrete or unnattended mode.  Discrete mode
is normally performed while a technician manages the Sonde during the
sampling process. In this mode, the Sonde is connected via 650
MDS Display/Logger or a serial cable connected to a PC, where the sampling
frequency is usually set to a high frequency. In contrast, Unattended
mode is performed when the Sonde is deployed offshore or in a
remote body of water, where the sampling frequency is usually set for a
longer period of time (i.e. 5 - 15 min). In order to be configured,
the Sonde offers a set of menus that can be accessed through a terminal session
over RS232. In this way, the Sonde can be configured to
start logging data to an internal file. In order to view the data as it's
being captured, the option to show live data is chosen and the data can then be 
seen via a serial terminal session at each interval.

@@

The DSP is implemented in Java and serves as a container environment where 
DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics, 
the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
infrastucture. One important DSP requirement is to impose on DSP components minimal 
dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
accessible to DSP components are specified by \texttt{DSPContext} interface.
The DSP also uses messages to communicate with DSP Components. It has a special "adapter"
that acts like a DSPComponent. Therefore, configuring the DSP is no different from configuring
a DSP component.

DSP compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
(bundles with DSP Components are transmitted as message to the DSP Component Adaptor) 
and activation/deactivation of DSP Components.

@@

In this paper, we introduce a practical application of a sensor network
in the domain of environmental monitoring. We have built an
environmental sensor network in the San Francisco Bay using highly
specialized, off-the-shelf water quality sensors. The resulting
infrastructure enables near-realtime access to sensor data collected
from remote locations. Our end-to-end solution, furthermore, allows
configuration and management of remotely deployed sensors. In this sense, the
key to our infrastructure is the DSP, a micro-kernel plug-and-play
architecture that can flexibly be configured to specific needs through
special purpose components.

>>>>

Each probe may have multiple sensors.  A computer can interface with
the Sonde via RS-232C or SDI-12. The Sonde can be configured to
collect data samples in discrete or unnattended mode.  Discrete mode
is normally performed while a technician manages the Sonde during the
sampling process. In this mode, the Sonde is connected via 650
MDS Display/Logger or a serial cable connected to a PC, where the sampling
frequency is usually set to a high frequency. In contrast, Unattended
mode is performed when the Sonde is deployed offshore or in a
remote body of water, where the sampling frequency is usually set for a
longer period of time (i.e. 5 - 15 min). Assuming a 5 min sampling
frequency, the Sonde produces 267 KB of data per day.
In order to be configured,
the Sonde offers a set of menus that can be accessed through a terminal session
over RS232. In this way, the Sonde can be configured to
start logging data to an internal file. In order to view the data as it's
being captured, the option to show live data is chosen and the data can then be 
seen via a serial terminal session at each interval.

@@

The DSP is implemented in Java and serves as a container environment where 
DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics, 
the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
infrastucture. One important DSP requirement is to impose on DSP Components minimal 
dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
accessible to DSP Components are specified by \texttt{DSPContext} interface.
The DSP also uses messages to communicate with DSP Components. It has a special adapter
that acts like a \texttt{DSPComponent}. Therefore, configuring the DSP is no different from configuring
a DSP Component.

DSP Compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
(bundles with DSP Components are transmitted as message to the DSP Component Adaptor) 
and activation/deactivation of DSP Components.

@@

In this paper, we introduce a practical application of a sensor network
in the domain of environmental monitoring. We have built an
environmental sensor network in the San Francisco Bay using highly
specialized, off-the-shelf water quality sensors. The resulting
infrastructure enables near-realtime access to sensor data collected
from remote locations. Our end-to-end solution, furthermore, allows
configuration and management of remotely deployed sensors. In this sense, the
key to our infrastructure is the DSP, a micro-kernel plug-and-play
architecture that can flexibly be configured to specific needs through
special purpose components. We plan to use the Gumstix's processing
power to pre-process data intensive sensors such as the HF Radar to
reduce the data volume.
